---
title: "Most streamed songs on Spotify in 2023"
output: html_notebook
---


```{r}
setwd("C:\\Users\\cjget\\Downloads")

data <- read.csv("C:\\Users\\cjget\\Downloads\\spotify-2023.csv")

View(spotify.2023)

```



1.) What are the top 5 most streamed songs streamed?
```{r}

# Sort the dataset by streams in descending order
sorted_data <- spotify.2023[order(-spotify.2023$streams), ]

# Select the top 5 rows
top_5_songs <- head(sorted_data, 5)

# Print the top 5 most streamed songs
print(top_5_songs)

```


2.) What was the valence and danceability of the #1 most streamed song?
```{r}
valence <- top_song$valence_
danceability <- top_song$danceability_

valence <- ifelse(is.na(valence), NA, as.numeric(valence))
danceability <- ifelse(is.na(danceability), NA, as.numeric(danceability))

# Print the valence and danceability values
cat("Valence of the top streamed song:", valence, "\n")
cat("Danceability of the top streamed song:", danceability, "\n")

# Create a bar plot for valence and danceability
barplot(c(valence, danceability), names.arg = c("Valence", "Danceability"), 
        col = c("blue", "green"), main = "Valence and Danceability of the Top Streamed Song",
        ylab = "Value", ylim = c(0, 100))


```

3.)How did Valence, energy and danceability affect the top 5 songs on the charts?
```{r}

spotify_clean <- spotify.2023[complete.cases(spotify.2023[, c("valence_.", "danceability_.", "energy_.", "streams")]), ]

# Get the top 5 songs based on streaming numbers
top_5_songs <- head(spotify_clean[order(-spotify_clean$streams), ], 5)

# Extract valence, danceability, and energy for the top 5 songs
valence <- top_5_songs$valence_.
danceability <- top_5_songs$danceability_.
energy <- top_5_songs$energy_.

# Calculate correlations between attributes and streaming numbers
cor_valence <- cor(top_5_songs$streams, valence)
cor_danceability <- cor(top_5_songs$streams, danceability)
cor_energy <- cor(top_5_songs$streams, energy)

# Print correlation coefficients
cat("Correlation between streaming numbers and valence:", cor_valence, "\n")
cat("Correlation between streaming numbers and danceability:", cor_danceability, "\n")
cat("Correlation between streaming numbers and energy:", cor_energy, "\n")

# Round correlation coefficients to 2 decimal places
cor_valence <- round(cor_valence, 2)
cor_danceability <- round(cor_danceability, 2)
cor_energy <- round(cor_energy, 2)

cat("Correlation between streaming numbers and valence:", cor_valence)
if (abs(cor_valence) >= 0.7) {
  cat(" (Strong correlation)")
} else if (abs(cor_valence) >= 0.3) {
  cat(" (Moderate correlation)")
} else {
  cat(" (Weak correlation)")
}
cat("\n")

cat("Correlation between streaming numbers and danceability:", cor_danceability)
if (abs(cor_danceability) >= 0.7) {
  cat(" (Strong correlation)")
} else if (abs(cor_danceability) >= 0.3) {
  cat(" (Moderate correlation)")
} else {
  cat(" (Weak correlation)")
}
cat("\n")

cat("Correlation between streaming numbers and energy:", cor_energy)
if (abs(cor_energy) >= 0.7) {
  cat(" (Strong correlation)")
} else if (abs(cor_energy) >= 0.3) {
  cat(" (Moderate correlation)")
} else {
  cat(" (Weak correlation)")
}
cat("\n")


```



4.) What artist had the most streams?
```{r}
spotify_clean <- spotify.2023[!is.na(spotify.2023$streams), ]

# Aggregate streams by artist
artist_streams <- aggregate(streams ~ artist.s._name, data = spotify_clean, sum)

# Find the artist with the most streams
top_artist <- artist_streams[which.max(artist_streams$streams), ]

# Print the artist with the most streams
cat("Artist with the most streams:", top_artist$artist.s._name, "\n")

```



5.) What were the top 3 songs that have the most energy?
```{r}
spotify_clean <- spotify.2023[!is.na(spotify.2023$energy_.), ]

# Sort the data by energy in descending order
sorted_data <- spotify_clean[order(-spotify_clean$energy_.), ]

# Select the top three rows
top_three_songs <- head(sorted_data, 3)

# Print the top three songs with the most energy
print(top_three_songs[, c("track_name", "energy_.")])

```



6.) How much danceability did the most streamed track in 2023 have?
```{r}

# Identify the most streamed track
most_streamed_track <- spotify.2023[which.max(spotify.2023$streams), ]

# Get the danceability of the most streamed track
most_streamed_danceability <- most_streamed_track$danceability

# Print the danceability of the most streamed track
cat("The danceability of the most streamed track is:", most_streamed_danceability, "\n")

```



7.) How well did the most streamed track do on spotify charts compared to other charts?
```{r}

# Remove rows with NA in the streams column
cleaned_spotify_2023 <- spotify.2023 %>%
  filter(!is.na(streams))

# Identify the most streamed track
most_streamed_track <- cleaned_spotify_2023[which.max(cleaned_spotify_2023$streams), ]

# Extract relevant information about the most streamed track
most_streamed_song_title <- most_streamed_track$song_title
most_streamed_track_charts <- most_streamed_track[, c("in_spotify_charts", "in_apple_charts", "in_deezer_charts", "in_shazam_charts")]

# Print the comparison of chart performance
cat("Chart performance for the most streamed track (", most_streamed_song_title, "):\n")
print(most_streamed_track_charts)

# Create a data frame for visualization
chart_data <- data.frame(
  Platform = c("Spotify", "Apple Music", "Deezer", "Shazam"),
  InCharts = c(
    most_streamed_track$in_spotify_charts,
    most_streamed_track$in_apple_charts,
    most_streamed_track$in_deezer_charts,
    most_streamed_track$in_shazam_charts
  )
)

# Create a bar chart
ggplot(chart_data, aes(x = Platform, y = InCharts, fill = Platform)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Chart Performance of the Most Streamed Track:", most_streamed_song_title),
       y = "Presence in Charts",
       x = "Platform") +
  theme_minimal()

```



8.) How does the valence (positveness) of a song relate to it's streaming numbers?
```{r}
 # Scatter plot for steams vs. valence

plot(spotify.2023$streams, spotify.2023$valence, main="Streams vs. Valence", xlab="Streams", ylab="Valence", col="skyblue")

options(scipen = 8000000000)
```




9.) How do streaming numbers vary by the release year of the songs? Are older songs just as popular or do the newer releases dominate the charts?
```{r}
# Convert 'streams' column to numeric
spotify.2023$streams <- as.numeric(as.character(spotify.2023$streams))

# Convert 'release_year' column to numeric
spotify.2023$release_year <- as.numeric(as.character(spotify.2023$release_year))

# Aggregate streams by release year
streams_by_year <- aggregate(streams ~ release_year, data = spotify.2023, sum)

options(scipen = 900000000)

# Plot streaming numbers by release year
plot(streams_by_year$release_year, streams_by_year$streams, type = "b", 
     main = "Streaming Numbers by Release Year", xlab = "Release Year", ylab = "Total Streams")

correlation <- cor(streams_by_year$release_year, streams_by_year$streams)

cat("Correlation coefficient between release year and total streams:", correlation, "\n")

# Compare popularity of older vs. newer songs
summary(streams_by_year$streams)


```



10.) Which song had the most liveness? Do songs with the most liveness get streamed more?
```{r}
# Find the row index of the song with the highest liveness
index_max_liveness <- which.max(spotify.2023$liveness)

# Get the name of the song with the highest liveness
song_max_liveness <- spotify.2023$track_name[index_max_liveness]

# Get the stream count of the song with the highest liveness
stream_count_max_liveness <- spotify.2023$streams[index_max_liveness]

# Print the information
cat("Song with the highest liveness:", song_max_liveness, "\n")
cat("Stream count of the song with the highest liveness:", stream_count_max_liveness, "\n")

# Analyze the relationship between liveness and streams
mean_stream_count_high_liveness <- mean(spotify.2023$streams[spotify.2023$liveness > 0.8], na.rm = TRUE)
cat("Mean stream count of songs with high liveness (>0.8):", mean_stream_count_high_liveness, "\n")

```



11.) How did collaborations affect song popularity?
```{r}

spotify_clean <- spotify.2023[complete.cases(spotify.2023[, c("artist_count", "streams")]), ]

summary_artist_count <- summary(spotify_clean$artist_count)
summary_streams <- summary(spotify_clean$streams)

# Print summary statistics
cat("Summary statistics for artist_count:", summary_artist_count, "\n")
cat("Summary statistics for streams:", summary_streams, "\n")

plot(spotify_clean$artist_count, spotify_clean$streams, 
     xlab = "Artist Count", ylab = "streams", 
     main = "Artist Count vs. streams",
     col = "blue", pch = 16, cex = 1.5, 
     xlim = c(0, max(spotify_clean$artist_count) + 1),
     ylim = c(0, max(spotify_clean$streams) + 1),
     cex.axis = 1.2, cex.lab = 1.2)

abline(lm(spotify_clean$streams ~ spotify_clean$artist_count), col = "red")


# Calculate correlation coefficient
correlation <- cor(spotify_clean$artist_count, spotify_clean$streams)
cat("Correlation between artist_count and streams:", correlation, "\n")

```


12.)
```{r}


```

